### 2014-08-23 Pero_Transcriptome.tex
this version sent to Mike

@@

To characterize natural selection on several genes related to water and ion homeostasis, we identified several of the transcripts identified as experiencing positive selection in a resent work on desert-adapter \textit{Dipodomys} rodents. The coding sequence corresponding to these genes, Solute Carrier family 2 member 9  (Slc2a9) and the Vitamin D3 receptor (Vdr), were extracted from the dataset, aligned using the software MACSE \cite{Ranwez:2011kj} to homologous sequences in \textit{Mus musculus}, \textit{Rattus norvegicus}, \textit{Peromyscus maniculatus}, and \textit{Homo sapens} identified by the conditional reciprocal best blast procedure (CRBB, \cite{Aubry:2014en}). An unrooted gene tree was constructed using the online resource Clustal-Omega, and together the tree and alignment were analyzed using the branch-site model (model=2, nsSites=2, fix\_omega=0 versus model=2, nsSites=2, fix\_omega=1, omega=1)  implemented in PAML version 4.8 \cite{Yang:2011bm,Yang:2007ki}. Significance was evaluated via use of the likelihood ratio test. \\

@@

As detailed above, the RNAseq data from 15 individuals were mapped to the reference transcriptome with the resulting BAM files being used as input to the software package ANGSD. The Tajima's D statistic was calculated for all  transcripts covered by at least 14 of the 15 individuals. The distribution of the results, shown in Figure 2, suggest that the vast majority of the transcriptome is under purifying selection (D \textless 0), with a much smaller fraction being subject to neutral or positive selection. \hyperlink{Table 3}{Table 3} presents the 10 transcripts whose estimate of Tajima's D is greatest, while \hyperlink{Table 4}{Table 4} presents the 10 transcripts whose estimate of Tajima's D is least. While the former list of genes is likely to contain transcripts experiencing positive selection in the studies population, the latter group is likely experiencing purifying selection. \\

@@

To test the hypothesis that selection on transcripts related to osmoregulation is enhanced in the desert adapted \peer, I implemented the branch-site test as described above, setting the sequence corresponding to \peer\:for both Slc2a9 and Vdr as the foreground lineages in 2 distinct program executions. The test for Slc2a9 was highly significant (2$\Delta$Lnl=51.4, df=1, p=0), indicating enhanced selection in \peer\: relative to the other lineages. The branch site test for positive selection conducted on the Vdr gene was non-significant (2$\Delta$Lnl=0.68, df=1, p=1).

@@

As a direct result of intense heat and aridity, deserts are thought to be amongst the most harsh of environments, particularly for it's mammalian inhabitants. Given the osmoregulation can be challenging for these animals, with failure resulting in death, strong selection should be observed on genes related to the maintenance of water and solute balance. This study aims to characterize the transcriptome of a desert-adapted rodent species, \peer. Specifically, we characterize the transcriptome of four tissue types (liver, kidney, brain, testes) from a single individual, and supplement this with population level renal transcriptome sequencing from 15 additional animals. This study aims to provide the foundation for future experimental research. \\

Previous studies of desert adaption in rodents have largely implicated the aquaporins \cite{Gallardo:2005fm,Huang:2001ti} in adaptation to desert environments (but see \cite{Marra:2012ec}), though most of these studies employed a candidate gene approach and therefore are limited in scope. The current study, which characterizes a multi-tissue transcriptome, nucleotide polymorphism, and natural selection

@@

MDM was supported by a NIH Postdoctoral fellowship, then by startup funds provided by the University of New Hampshire. He would like to acknowledge the support of his wife and children. Lastly, this work was significantly improved by comments received based on a version of the manuscript posted on biorxiv.

>>>>

As a direct result of intense heat and aridity, deserts are thought to be amongst the most harsh of environments, particularly for it's mammalian inhabitants. Given the osmoregulation can be challenging for these animals, with failure resulting in death, strong selection should be observed on genes related to the maintenance of water and solute balance. This study aims to characterize the transcriptome of a desert-adapted rodent species, \textit{Peromyscus eremicus}. Specifically, we characterized the transcriptome of four tissue types (liver, kidney, brain, testes) from a single individual, and supplement this with population level renal transcriptome sequencing from 15 additional animals. We identified a set of transcripts undergoing both purifying and balancing selection based on estimates of Tajima's D. In addition, we used a branch site test to identify a transcript, Slc2a9 likely related to desert osmoregulation, undergoing enhanced selection in \peer\: relative to a set of non-desert rodents.  \\

@@

To characterize natural selection on several genes related to water and ion homeostasis, we identified several of the transcripts identified as experiencing positive selection in a resent work on desert-adapter \textit{Dipodomys} rodents. The coding sequence corresponding to these genes, Solute Carrier family 2 member 9  (Slc2a9) and the Vitamin D3 receptor (Vdr), were extracted from the dataset, aligned using the software MACSE \cite{Ranwez:2011kj} to homologous sequences in \textit{Mus musculus}, \textit{Rattus norvegicus}, \textit{Peromyscus maniculatus}, and \textit{Homo sapiens} identified by the conditional reciprocal best blast procedure (CRBB, \cite{Aubry:2014en}). An unrooted gene tree was constructed using the online resource Clustal-Omega, and together the tree and alignment were analyzed using the branch-site model (model=2, nsSites=2, fix\_omega=0 versus model=2, nsSites=2, fix\_omega=1, omega=1)  implemented in PAML version 4.8 \cite{Yang:2011bm,Yang:2007ki}. Significance was evaluated via use of the likelihood ratio test. \\

@@

As detailed above, the RNAseq data from 15 individuals were mapped to the reference transcriptome with the resulting BAM files being used as input to the software package ANGSD. The Tajima's D statistic was calculated for all  transcripts covered by at least 14 of the 15 individuals. In brief, a negative Tajima's D, a result of lower than expected average heterozygosity, is often associated with purifying or directional selection, recent selective sweep or population bottleneck. In contrast, a positive value for Tajima's D represents higher that expected average heterozygosity, often associated with balancing selection. \\

The distribution of the estimates of Tajima's D for all assembled transcripts is shown in Figure 2. The distribution is skewed towards negative values (mean=-0.89, variance=0.58), which is likely the result of purifying selection, a model of evolution commonly invoked for coding DNA sequences \cite{Chamary:2006db}. \hyperlink{Table 3}{Table 3} presents the 10 transcripts whose estimate of Tajima's D is greatest, while \hyperlink{Table 4}{Table 4} presents the 10 transcripts whose estimate of Tajima's D is least. The former list of genes is likely to contain transcripts experiencing positive selection in the studied population. This list includes, interestingly, genes obviously related to solute and water balance (e.g. Clcnkb and a transmembrane protein gene), and those related to immune function (a interferon-inducible GTPase and a Class 1 MHC gene). The latter group, containing the transcripts whose estimates of Tajima's D is the least are likely experiencing purifying selection. Many of these transcripts are involved in core regulatory functions where mutation may have strongly negative fitness consequences. \\

@@

To begin to test the hypothesis that selection on transcripts related to osmoregulation is enhanced in the desert adapted \peer, I implemented the branch-site test as described above, setting the sequence corresponding to \peer\:for both Slc2a9 and Vdr as the foreground lineages in 2 distinct program executions. These two transcripts were chose specifically because they, the former significantly, were recently linked to osmoregulation in a desert rodent \cite{Marra:2014de}. The test for Slc2a9 was highly significant (2$\Delta$Lnl=51.4, df=1, p=0), indicating enhanced selection in \peer\: relative to the other lineages. The branch site test for positive selection conducted on the Vdr gene was non-significant (2$\Delta$Lnl=0.68, df=1, p=1). This limited analysis of selection is to be followed up by an analysis of genome wide patterns of natural selection. \\

@@

As a direct result of intense heat and aridity, deserts are thought to be amongst the most harsh of environments, particularly for it's mammalian inhabitants. Given the osmoregulation can be challenging for these animals, with failure resulting in death, strong selection should be observed on genes related to the maintenance of water and solute balance. This study aimed to characterize the transcriptome of a desert-adapted rodent species, \peer. Specifically, we characterized the transcriptome of four tissue types (liver, kidney, brain, testes) from a single individual, and supplement this with population level renal transcriptome sequencing from 15 additional animals. We identified a set of transcripts undergoing both purifying and balancing selection based on Tajima's D. In addition, we used a branch site test to identify a transcript, likely related to desert osmoregulation, undergoing enhanced selection in \peer\: relative to a set of non-desert rodents. This study aims to provide the foundation for future experimental research, attempting to understand the genetic underpinnings of extreme osmoregulation in  desert rodents. \\

@@

MDM would like to acknowledge the support of his wife and children. Lastly, this work was significantly improved by comments received based on a version of the manuscript posted on biorxiv.
